### 2018-06-12 99_Documentation/diogo_perso.tex
New version of Paper : V4 (plus little fixes at various places)

\begin{tikzpicture}
\tikzset{every node/.style={on chain,draw,thick,rounded corners,
minimum height=3em, text width=10em, align=center}}
\begin{scope}[start chain=going below]
\node (prepare-group) {Prepare group list (\texttt{.csv})};
\node (import-group) {Import group list for new score file (\texttt{.xml})};
\node (enter-rating) {Enter peer ratings given by each student};
\node (compute-score) {Compute autorated weights and scores};
\node (export-csv) {Export data for Excel (\texttt{.csv})};
\end{scope}
\node[right=of enter-rating] (open-score) {Open previously saved score file (\texttt{.xml})};

\path[draw,line width=0.4ex, ->,>= angle 60]
(prepare-group) edge (import-group) 
(import-group) edge (enter-rating)
(open-score) edge (enter-rating)
(enter-rating) edge (compute-score) 
(compute-score) edge (export-csv);
\end{tikzpicture}