### 2018-06-25 99_Documentation/14_StabFem_plotting.tex

% !TEX root = main.tex

\chapter{Plotting data with StabFem}

\section{Plotting with StabFem : \texttt{plotFF.m}}

\section{Native FreeFem plots (ffglut)}

this is enabled/disabled with the MACRO FREEFEMPLOTS

\section{Interfacing with Tecplot}

\texttt{exportFF\_Tecplot.m}

\section{Interfacing with Paraview}

Javier did that ???

\section{Alternative idea}

The problem with plotFF is that is is built upon pdeplot and requires the pdetools toolbox.

An alternative solution would be to adapt the solution of chloros based on patch, which has the advantage of operating under OCTAVE.

See more there :  \verb| https://github.com/samplemaker/freefem_matlab_octave_plot  |