### 2018-07-17 99_Documentation/MANUAL/03_Previous_Note.tex

\section*{Previous Note}

\ti{StabFem} is a software to perform Global Stability calculations in Fluid Mechanics, which is developed for both research and education purposes. It can be downloaded at \url{https://github.com/erbafdavid/StabFem}. Additionally, one will previously need to install \ti{Matlab} and \ti{FreeFem++} on the running system. For now, \ti{StabFem} works perfectly in \ti{Ubuntu16.06}. On other Operating Systems, it's not sure that it will work.

This document is a collaboration of different people and is intended to explain, the best way possible, the different parts of \textit{StabFem}. Perhaps as the reader will read it, some chapters are uncompleted or complicated to understand. The reader is invited to contribute in order to render the documentation more understandable.