### 2018-07-17 99_Documentation/MANUAL/301_Mutual_Files_Matlab.tex

% !TEX root = main.tex
\chapter{Mutual Matlab Files}

In this section we give the documentation of the Matlab drivers of the project.

We have to find a way to generate automatically this documentation by extracting the documentation directly from the programs
(Doxygen ?)

\section{General drivers and FF/Matlab interface tools}

\subsection{\texttt{SF\_Mesh.m}}

\subsection{\texttt{SF\_Launch.m}}

\subsection{\texttt{importFFMesh.m}}

\subsection{\texttt{importFFData.m}}

\section{Stability-oriented drivers}

\subsection{\texttt{SF\_Init.m}}

\subsection{\texttt{SF\_BaseFlow.m}}

\subsection{\texttt{SF\_Stability.m}}

\subsection{\texttt{SF\_Adapt.m}}

\section{Plotting programs}

\section{Auxiliary programs}

\subsection{\texttt{mysystem.m}}

\subsection{\texttt{mycp.m}}