### 2018-07-17 99_Documentation/MANUAL/04_General_view.tex

\chapter{General View of \ti{StabFem}}

\ti{StabFem} has a main directory where all the projects are located and where the mutual scripts are located too. The directory is composed by the following particular project directories:

\begin{enumerate}
\item ACOUSTICS\_PIPES;
\item CYLINDER;
\item CYLINDER\_VIV;
\item DISK\_IN\_PIPE.
\item etc...
\end{enumerate}
The mutual directories are the followings:

\begin{enumerate}
\item SOURCES\_FREEFEM;
\item SOURCES\_MATLAB;
\item Documentation.
\end{enumerate}

These directories are used by the several projects of \ti{StabFem}.

\medskip

\begin{leftbar}
Attention: When you do some change in the mutual directory files, we have to assure that it will work on all the projects.
\end{leftbar}
\section{General Features}

\begin{itemize}
\item With \ti{StabFem} we can solve forced problems, eigenvalue problems, etc...

\item  \ti{StabFem} allows us to save the scripts used to generate the different data and graphs, unabling to reproduce the same results (useful for repeat figures for articles, confirming results, etc...).
\end{itemize}

\section{What is attended from this manual document}
This document presents this first chapter describing what can be found in a  \ti{StabFem} project, a chapter describing the running logic of a \ti{StabFem} program and the main common files and a chapter for each project, describing succinctly their running features.