### 2018-07-20 99_Documentation/MANUAL/212_SessileDrops.tex
commit sur branche nabil

% !TEX root = main.tex

\chapter{Linear dynamics of sessile drops}

\begin{description}
\item{Directory in the StabFem project:}  \texttt{SessileDropsBostwickSteen}
\item{Main contributors:} Nabil ACHOUR \& David FABRE.
\item{Reference:} Linear and nonlinear oscillations of sessile drops and partially filled container, N. Achour.  
\end{description}

A sessile droplet partially wets a planar solid support. We study the linear stability of an inviscid sessile drops. In this test-case the three-phase contact line is pinned.

\section{Creation of initial mesh for Sessile drops}

\section{Eigenvalue computation for m=0 and m=1}